---
title: "Practical Machine Learning Project"
author: "Iris Guan"
output: html_document
---
## Project Goal

The goal of this project is to predict the manner in which people did the exercise. 

# Data Download and Preperation

```{r dataprep, echo=TRUE, message=FALSE, warning=FALSE}
        set.seed(1)
        
        library(mlbench)
        library(caret)

        download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile = "./training.csv")
        training <- read.csv("./training.csv", header=TRUE,na.strings=c("NA","#DIV/0!", ""))

        download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile= "./testing.csv")
        testing <- read.csv("./testing.csv", header=TRUE, na.strings=c("NA","#DIV/0!", ""))
```
#Remove features that don't help on prediction or show low vairance, from training data. Also remove features that have more than 80% NA values.

```{r fct, echo=TRUE, message=FALSE, warning=FALSE}
        FC <- grep("X|timestamp|name|window", colnames(training))
        training <- training[,-FC]
        nearzero <- nearZeroVar(training, saveMetrics = TRUE)
        training <- training[, !nearzero$nzv]

        Index80NA <- sapply(colnames(training), function(x) if(sum(is.na(training[, x])) > 0.80*nrow(training)) {return(TRUE)} else{return(FALSE)})
        training <- training[, !Index80NA]

        dim(training)
```

#Check correlation among remaining features

```{r corcheck, echo=TRUE, message=FALSE, warning=FALSE}
        correlationMatrix <- cor(training[,-53], use="pairwise.complete.obs") 
        highlyCorrelated <- findCorrelation(correlationMatrix,cutoff=0.75)
        names(training)[highlyCorrelated]
```
#We could see that many variables are correlated. We will apply PCA analysis to reduce feature dimension on subset of trainign data for cross validation. 
#Cross Vadalidation Data Setup

```{r cv, echo=TRUE, message=FALSE, warning=FALSE}
        dataInx <- createDataPartition(training$classe, p=0.75, list=F)

        Ttrain <- training[dataInx,]
        Ttest <- training[-dataInx,]
```

#PCA Analysis on training subset

```{r pca, echo=TRUE, message=FALSE, warning=FALSE}
        pPCA <- preProcess(Ttrain, method="pca", thresh=0.75)
        TrainData <- predict(pPCA, Ttrain)
```

#Random Forest

```{r rf, echo=TRUE, message=FALSE, warning=FALSE}
        library(randomForest)
        NAclasseInx <- is.na(TrainData$classe)
        TrainData <- TrainData[!NAclasseInx,]
        RFModel <- randomForest(classe ~ ., data=TrainData, do.trace=F)
        print(RFModel)
```

#Apply on test data for validation

```{r tdcv, echo=TRUE, message=FALSE, warning=FALSE}
        TestData <- predict(pPCA, Ttest)
        TestDataRF <- predict(RFModel, TestData)
        confusionMatrix(Ttest$classe,TestDataRF)$overall[["Accuracy"]]
```

#Apply on the 20 test cases available in the test data

```{r tds, echo=TRUE, message=FALSE, warning=FALSE}
        testing <- testing[,-FC]

        Tnearzero <- nearZeroVar(testing, saveMetrics = TRUE)
        testing <- testing[, !Tnearzero$nzv]

        TIndex80NA <- sapply(colnames(testing), function(x) if(sum(is.na(testing[, x])) > 0.80*nrow(testing)) {return(TRUE)} else{return(FALSE)})
        testing <- testing[, !TIndex80NA]

        T20 <- predict(pPCA, testing)
        testing$classe <- predict(RFModel, T20)
        
        testing$classe

```
